Supplementary Materials to ``Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic''

### 2022-09-19 write-up/letters/response-ref-1-v1.tex

\item {\it On the plots that show the active infection rate estimates, perhaps including a line that is the case counts divided by the population as a floor would be worthwhile (at least in the supplement) to demonstrate that the model-based estimates are at least above this.}

\vspace{5mm}
We have added the floor of weekly case counts divided by population size to the figure of active infection rate estimates.  We interpret ``case counts'' in the referee comment above as ``recent case counts'' so that the floor calculation is relevant for estimating the active infection rate.
\vspace{5mm}

\item {\it I found the grey hard to differentiate for the figures that have more than two categories – perhaps varying the shapes would help if trying to keep these grey-scale?}

\vspace{5mm}
We tested a variety of means to differentiate among categories in figures with more than two categories.  We found varied shapes made the plots much harder to understand in settings with significant overlap.  Therefore, we opted to simply remove the gray background.  The white background helps with differentiation while keeping the figures in grey-scale to match AOAS figure guidelines.
\vspace{5mm}

### 2022-09-19 write-up/letters/response-ref-1-v1.tex

\item {\it The reliance on the random sample with such a low response rate may be subject to unmeasured confounding - perhaps a sensitivity analysis showing the impact of a potential unmeasured confounder could strengthen this point.}

\vspace{5mm}
We have added a sensitivity analysis in Section K of the supplementary materials. Note the difficulty of such an analysis comes from needing to specify a joint distribution of the measured covariates~$X_t$, the potential unmeasured and \emph{time-varying} confounder~$U_t$, the self-selection indicator~$I_t$ and the outcome~$Y_t$.  Here, we extend recent work in the causal inference literature on sensitivity analysis to the self-selection setting in order to avoid direct specification of these distributions.  Instead, we specify a conditional distribution of the self-selection propensity~$\pi_t(X_t, U_t)$ given the estimated propensity~$\pi_t (X_t)$ and use this as the basis for a sensitivity analysis to derive a simple expression for potential bias.  The bias depends on two sensitivity parameters: (1) $\alpha_t$ which measures the confounder-selection relationship strength, and (2) $\delta_t$ which measures the confounder-outcome relationship strength.  We calibrate the selected sensitivity parameter values by analyzing the observed data as if a particular observed covariate were an unobserved confounder.  A complete set of derivations and associated discussion are presented in Section K of the supplementary materials, which we briefly discuss in Section 5.3 of the main manuscript.
\vspace{5mm}

\item {\it On the plots that show the active infection rate estimates, perhaps including a line that is the case counts divided by the population as a floor would be worthwhile (at least in the supplement) to demonstrate that the model-based estimates are at least above this.}

\vspace{5mm}
We have added the floor of weekly case counts divided by population size to Figure 11 in the supplementary materials.  We interpret ``case counts'' in the referee comment above as ``weekly case counts'' so that the floor calculation is relevant for estimating the active infection rate.  Note all estimates are above the floor with the doubly robust being closest to the floor in September.
\vspace{5mm}

### 2022-09-20 write-up/letters/cover letter-covid.tex
Cov Letter added
Updated papers - almost there

\begin{letter}{Professor Karen Kafadar \\
    Editor-In-Chief, {\em Annals of Applied Statistics}
         }
%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

@@

\opening{Dear Dr. Kafadar:}

Kindly find enclosed the revised version of the manuscript number AOAS2006-048 titled ``Statistical paradoxes in coronavirus case-counts: selection bias, measurement error, and the COVID-19 pandemic'' now titled ``Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information.''  We have revised the article to address the referees’ requests about exposition and other points of clarification.  The reports called for a major revision to make the manuscript ``forward compatible'', to re-structure the manuscript to present figures and tables near their corresponding discussion, to discuss potential non-response bias in the motivating studies, to include a more careful discussion of COVID-19 testing and disease dynamics, to include a more precise definition of the active infection rate and how it impacts model-based estimation, and to more carefully tie the proposal to related parts of the epidemiological literature. The new version complies with these requests.

We believe the work presents an important advance in how researchers should think about addressing time-varying selection bias and measurement error to better understanding the COVID-19 pandemic trajectory. Attached please find a detailed response to the referees outlining how the changes to the manuscript address the concerns raised in their reports.

>>>>

\begin{letter}{Professor Ji Zhu \\
    Editor-In-Chief, {\em Annals of Applied Statistics}
         }
%----------------------------------------------------------------------------------------
%	YOUR NAME & ADDRESS SECTION
%----------------------------------------------------------------------------------------

@@

\opening{Dear Dr. Zhu:}

Kindly find enclosed the revised version of the manuscript number AOAS2006-048 titled ``Addressing selection bias and measurement error in COVID-19 case count data using auxiliary information.''  We are happy that the two reviewers, associate editor, and editor have all found the revision much improved. We have revised the article to address the few remaining concerns that one of the reviewers raised but are glad that they ``appreciate the paper and [are] very supportive of the pragmatic approach.''  We have edited the remark, performed a sensitivy analysis to assess sensitivity due to potential confounding, and adjusted the figures to assist in comprehension.

We continue to believe the work presents an important advance in how researchers should think about addressing time-varying selection bias and measurement error to better understanding the COVID-19 pandemic trajectory. Attached please find a detailed response to the referee outlining how the changes to the manuscript address the concerns raised in their report.
